Exploring the Design Space of Differentially Private Bayesian Inference under Hellinger Distance

### 2019-01-22 ICML2019/main.tex
work on 5

\section{Private Mechanisms}
\label{sec_mechs}
In this section we will describe and analyse a series of mechanisms.
Section \ref{sec_ell1_dim} presents a first mechanism, $\lapmech$,
based on the Laplace primitive where the sensitivity is measured w.r.t. to $\ell_1$ norm.
Section \ref{sec_ell1_hist}, shows that $\lapmech$'s accuracy can be improved
by performing a better analysis on the sensitivity of the inference process.
Section \ref{sec_ehd}, shows an instance of the exponential mechanism, $\expmech$, where the scoring
function is based on the Hellinger distance, we will see how this mechanisms performs
poorly due to the very high global sensitivity of the inference process w.r.t to this metric.
Section \ref{sec_ehdl} introduces, $\lexpmech$, a variation of $\expmech$ which uses a lower bound
on the global sensitivity called local sensitivity.
Unfortunately, despite improving on accuracy $\lexpmech$ is non-private.
Finally, building on $\lexpmech$, Section \ref{sec_ehds} introduces $\hexpmech$:
which, based on the notion of \emph{smooth} sensitivity, achieves $\epsilon$-differential privacy and
has good accuracy in specific cases.
In presenting these mechanisms we will denote by $\inferd{\valpha}{\vtheta}{\dataobs}$ the posterior  $\dirichlet{\valpha'}$
obtained starting from the prior $\dirichlet{\valpha}$, likelihood $\multinomial{|\dataobs|}{\vtheta}$, and observed data $\dataobs$.
Two different upper bounds to the value
\begin{align}
  \label{eqn:eq1}
  \max_{\substack{\adj{\dataobs}{\dataobs'},\\
      \dirichlet{\valpha}=\inferd{\valpha}{\vtheta}{\dataobs},\\
      \dirichlet{\valpha'}=\inferd{\valpha}{\vtheta}{\dataobs'}}}||\valpha-\valpha'||_1 
\end{align}
will be used to calibrate the noise in the mechanisms based on the Laplacian primitive.

\subsection{Calibrating noise w.r.t. $\ell_1$ norm ($\lapmech$)}
\label{sec_ell1_dim}
We start off by presenting $\lapmech$, in Algorithm \ref{lapmech}, based on the Laplace primitive.
The idea of $\lapmech$ is to add Laplacian noise to the differences of the respective numeric parameters of the prior
and posterior distribution. 
\begin{algorithm}
  \caption{$\lapmech$}
  \label{lapmech}
  \begin{algorithmic}
  \INPUT $\dataobs\in\mathcal{X}^n$, $\dirichlet{\valpha}$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \inferd{\valpha}{\vtheta}{\dataobs}$
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )\in \mathbb{N}^{|\mathcal{X}|}$ 
  \STATE \quad {\bf For} $i = 1 \dots |\mathcal{X}| - 1$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{|\mathcal{X}|}{\epsilon}}$
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_{|\mathcal{X}|}} = \alpha_{|\mathcal{X}|} + \lfloor n - \sum_{i = 1}^{|\mathcal{X}|-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\tilde{\valpha}$
  \end{algorithmic}
\end{algorithm}
\begin{lem}
  Algorithm \ref{lapmech} is $\epsilon$-d.p.
  \begin{proof} By noticing that $\mathcal{X}-1$ is an upper bound to \ref{eqn:eq1}. \end{proof}
\end{lem}

\subsection{Improving the sensitivity ($\ilapmech$)}
\label{sec_ell1_hist}
By seeing the parameters of the distributions as histograms, we can give a tighter upper bound to \ref{eqn:eq1}.
Indeed  $\lapmech$ can be reduced to a constant. Algorithm $\ilapmech$ in Algorithm \ref{mech:ilapmech} takes advantage of this fact. 
Indeed, when $\mathcal{X}=2$, \ref{eqn:eq1} is upper bounded by 1, while when $\mathcal{X}=2$ \ref{eqn:eq1} is upper bounded by 2.
In Algorithm \ref{mech:ilapmech} the algorithm $\ilapmech$ is presented.
\begin{algorithm}
  \caption{$\ilapmech$ in Dirichlet-Multinomial model}
  \label{mech:ilapmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\dirichlet{\valpha}$, likelihood: $\multinomial{n}{\vtheta}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \bysinfer(\multinomial{n}{\vtheta}, \dirichlet{\valpha}, \dataobs)$, $k = |\mathcal{X}|$.
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )$ of $k$ dimensions .
  \STATE \quad {\bf For} $i = 1 \dots (k - 1)$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_k} = \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\dirichlet{\tilde{\valpha}}$
  \end{algorithmic}
  \end{algorithm}

>>>>

\section{Private Mechanisms}
\label{sec_mechs}
In this section we will describe and analyse a series of mechanisms.
Section \ref{sec_ell1_dim} presents a first mechanism, $\lapmech$,
based on the Laplace primitive where the sensitivity is measured w.r.t. to $\ell_1$ norm.
Section \ref{sec_ell1_hist}, shows that $\lapmech$'s accuracy can be improved
by performing a better analysis on the sensitivity of the inference process.
Section \ref{sec_ehd}, shows an instance of the exponential mechanism, $\expmech$, where the scoring
function is based on the Hellinger distance, we will see how this mechanisms performs
poorly due to the very high global sensitivity of the inference process w.r.t to this metric.
Section \ref{sec_ehdl} introduces, $\lexpmech$, a variation of $\expmech$ which uses a lower bound
on the global sensitivity called local sensitivity.
Unfortunately, despite improving on accuracy $\lexpmech$ is non-private.
Finally, building on $\lexpmech$, Section \ref{sec_ehds} introduces $\hexpmech$:
which, based on the notion of \emph{smooth} sensitivity, achieves $\epsilon$-differential privacy and
has good accuracy in specific cases.
In presenting these mechanisms we will denote by $\inferd{\valpha}{\vtheta}{\dataobs}$ the posterior  $\dirichlet{\valpha'}$
obtained starting from the prior $\dirichlet{\valpha}$, likelihood $\multinomial{|\dataobs|}{\vtheta}$, and observed data $\dataobs$.
Two different upper bounds to the value of
\begin{align}
  \label{eqn:eq1}
  \max_{\substack{\adj{\dataobs}{\dataobs'},\\
      \dirichlet{\valpha}=\inferd{\valpha}{\vtheta}{\dataobs},\\
      \dirichlet{\valpha'}=\inferd{\valpha}{\vtheta}{\dataobs'}}}||\valpha-\valpha'||_1 
\end{align}
will be used to calibrate the noise in the mechanisms based on the Laplacian primitive.

\subsection{Calibrating noise w.r.t. $\ell_1$ norm ($\lapmech$)}
\label{sec_ell1_dim}
We start off by presenting $\lapmech$, in Algorithm \ref{lapmech}, based on the Laplace primitive.
The idea of $\lapmech$ is to add Laplacian noise with scale proportional to $\mathcal{X}$ to
the differences between the respective numeric parameters of the prior
and posterior distribution. 
\begin{algorithm}
  \caption{$\lapmech$}
  \label{lapmech}
  \begin{algorithmic}
  \INPUT $\dataobs\in\mathcal{X}^n$, $\dirichlet{\valpha}$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \inferd{\valpha}{\vtheta}{\dataobs}$
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )\in \mathbb{N}^{|\mathcal{X}|}$ 
  \STATE \quad {\bf For} $i = 1 \dots |\mathcal{X}| - 1$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{|\mathcal{X}|}{\epsilon}}$
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_{|\mathcal{X}|}} = \alpha_{|\mathcal{X}|} + \lfloor n - \sum_{i = 1}^{|\mathcal{X}|-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\tilde{\valpha}$
  \end{algorithmic}
\end{algorithm}
\begin{lem}
  Algorithm \ref{lapmech} is $\epsilon$-d.p.
  \begin{proof} By noticing that $\mathcal{X}-1$ is an upper bound to the value denoted by (\ref{eqn:eq1}). \end{proof}
\end{lem}

\subsection{Improving the sensitivity ($\ilapmech$)}
\label{sec_ell1_hist}
By seeing the parameters of the distributions as histograms, we can give a tighter upper bound to the value denoted by (\ref{eqn:eq1}).
Indeed a constant will do. Algorithm $\ilapmech$ in Algorithm \ref{ilapmech} takes advantage of this fact. 
Indeed, when $\mathcal{X}=2$, 1 is an upper bound for the value of (\ref{eqn:eq1}), while when $\mathcal{X}>2$, 2 is.
In Algorithm \ref{ilapmech} the algorithm $\ilapmech$ is presented.
\begin{algorithm}
  \caption{$\ilapmech$}
  \label{ilapmech}
  \begin{algorithmic}
  \INPUT $\dataobs\in\mathcal{X}^n$, $\dirichlet{\valpha}$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \inferd{\valpha}{\vtheta}{\dataobs}$
  \STATE \quad ${\bf let}\ k=\left \{
        \begin{array}{lll}          
          1&\text{if}& |\mathcal{X}|=2\\
          2&\text{otherwise}
        \end{array}
      \right .\ $
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )\in \mathbb{N}^{|\mathcal{X}|}$ 
  \STATE \quad {\bf For} $i = 1 \dots |\mathcal{X}|-1$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{k}{\epsilon}}$
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_{|\mathcal{X}|}} = \alpha_{|\mathcal{X}|} + \lfloor n - \sum_{i = 1}^{|\mathcal{X}|-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\tilde{\valpha}$
  \end{algorithmic}
\end{algorithm}
\begin{lem}
  Algorithm \ref{ilapmech} is $\epsilon$-d.p.
  \begin{proof} By noticing that 1 and 2 are upper bounds to the value denoted by (\ref{eqn:eq1}), respectively when $|\mathcal{X}|=2$ and $|\mathcal{X}|>2$.
  \end{proof}
\end{lem}
